---
title: "ps1_q3"
author: "Jingxian(Derrick) Chen"
date: "Sep 20th, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I have tried to apply the data from Stats506_F19 repo on Github to check my solutions
against the sample measures.

I will display my functions and results in the order of the questions and show 
the formatted table for the metrics for the test trajectories at last.

---

# Get sample data
First, I need to process the sample data in order to use in my functions.  
Here I choose part of the train_trajectories data giving subject_nr is 1 and count_trial is 2.  

*subset* will be an n*3 matrix representng the trajectory (x, y, t) giving 
subject_nr is 1 and count_trial is 2.
```{r import data, echo=FALSE, warning=FALSE, message=FALSE}
library(readr)
train_trajectories <- read_csv("C:/Users/95260/Desktop/git/jbhender/Stats506_F19/problem_sets/data/train_trajectories.csv")
test_trajectories <- read_csv("C:/Users/95260/Desktop/git/jbhender/Stats506_F19/problem_sets/data/test_trajectories.csv")
train_measures <- read_csv("C:/Users/95260/Desktop/git/jbhender/Stats506_F19/problem_sets/data/train_measures.csv")
```

### **Relevant code:**
```{r sample data}
subset <- as.matrix(train_trajectories[train_trajectories$subject_nr == 1 
        & train_trajectories$count_trial == 2, 3:5])
head(subset,3)
```

---

# Solution (a)
Functon description: to solve question 3a, translate input data to begin with 
time zero at the origin.  

Input: an n * 3 matrix representing the trajectory (x, y, t).

Output: an n * 3 matrix representing the trajectory (x, y, t) beginning with time zero at the origin.

### **Function code:**
```{r translate}
translate <- function(trajectories_data) {
  translate_data <- trajectories_data
  translate_data[, 1] <- trajectories_data[, 1] - trajectories_data[1, 1]
  translate_data[, 2] <- trajectories_data[, 2] - trajectories_data[1, 2]
  translate_data[, 3] <- trajectories_data[, 3] - trajectories_data[1, 3]
  translate_data
}
trans_subset <- translate(subset)
head(trans_subset,3)
```

We could see that the sample data was successfully translated to begin with time
zero at the origin.

---

# Solution (b)
Function description: to compute the angle $\theta$ formed by the secant line connecting the origin and the final position in the trajectory.

Input: the translated n * 3 matrix representing the trajectory (x, y, t).

Output: a double digit representing the angle between $[-\pi, \pi]$.

### **Function code**
```{r angle_compute}
angle_compute <- function(trajectories_data) {
  x_final <- trajectories_data[nrow(trajectories_data), 1]
  y_final <- trajectories_data[nrow(trajectories_data), 2]
  angle <- as.double(atan2(y_final, x_final))
  angle
}
theta <- angle_compute(trans_subset)
theta
```

So I have successfully calculated that the angle for the sample data.

---

# Solution (c)
Function description: to rotate the (x, y) coordinates of a trajectory so that the final point lies along the positive x-axis.

Input:   
$\# 1$: the translated n * 3 matrix representing the trajectory (x, y, t).  
$\# 2$: the angle computed by the angle_compute function.

Output: an n * 3 matrix representing the trajectory (x, y, t) whose final point
lies along the positive x-axis. 

### **Function code**
```{r rotate}
rotate <- function(trajectories_data, theta) {
# A represents the rotate matrix.
  new_data <- trajectories_data
  A <- matrix(c(cos(theta), -sin(theta), sin(theta), cos(theta)), nrow = 2)
  new_data[, 1:2] <- t(A %*% t(trajectories_data[, 1:2]))
  new_data
}
Rotate_subset <- rotate(trans_subset,theta)
head(Rotate_subset,3)
tail(Rotate_subset,3)
```

So I have successfully rotated the sample data to begin at the origin and let its
final point lies along the positive x-axis. 

---

# Solution (d)
Function description: to combine the functions above that normalizes an n * 3
trajectory matrix to begin at the origin and end on the positive x-axis.

Input: the origin n * 3 matrix representing the trajectory (x, y, t)

Output: an n * 3 trajectory matrix to begin at the origin and end on the positive x-axis.

### **Function code**
```{r normalize}
normalize <- function(trajectories_data) {
  translate_data <- translate(trajectories_data)
  theta <- angle_compute(translate_data)
  normalize_data <- rotate(translate_data, theta)
  normalize_data
}
normalize_subset <- normalize(subset)
head(normalize_subset,3)
tail(normalize_subset,3)
```

So, I have successfully got the same result from normalize function and the first three functions.

---

# Solution (e)
Function description: to compute the values of question e giving specific 
subject_nr and count_trial.

Input: the normalized trajectory matrix.

Output: a 1 * 4 vector consists of total (Euclidean) distance traveled, maximum
absolute deviation, average absolute deviation, absolute area.

Reference formulas:

(1) $$tot\_dist = \sum_{i=1}^{n-1}\sqrt{(x_{i}-x_{i+1})^2+(y_{i}-y_{i+1})^2}$$

(2) $$max\_abs\_dev = max_i{|\frac{(y_{n}*x_i-x_n*y_i)}{\sqrt{y_n^2+x_n^2}}|}$$

(3) $$avg\_abs\_dev = \frac{1}{n}*\sum_{i=1}^{n}|\frac{(y_{n}*x_i-x_n*y_i)}{\sqrt{y_n^2+x_n^2}}|$$

(4) $$AUC = \sum_{i=1}^{n-1}(x_{i+1}-x_i)*\frac{|y_i|+|y_{i+1}|}{2}$$

### **Function code**
```{r compute}
compute <- function(trajectories_data) {
  data <- normalize(trajectories_data)
  tot_dist <- 0
  max_abs_dev <- 0
# temp_abs_dev is for computing the max_abs_dev.
  temp_abs_dev <- 0
  avg_abs_dev <- 0
  AUC <- 0
  t_final <- data[nrow(data), 3]
  x_final <- data[nrow(data), 1]
  y_final <- data[nrow(data), 2]
# d2 is for computing the temporary absolute deviation.
  d2 <- (y_final)^2 + (x_final)^2
  for (i in 1:nrow(data)) {
    temp_abs_dev <- abs(y_final * data[i, 1] - x_final * data[i, 2]) / sqrt(d2)
    avg_abs_dev <- avg_abs_dev + temp_abs_dev
    if (i != nrow(data)) {
# d1 is for computing the total distance.
      d1 <- (data[i, 1] - data[i + 1, 1])^2 + (data[i, 2] - data[i + 1, 2])^2
      tot_dist <- tot_dist + sqrt(d1)
      AUC <- AUC + diff(data[, 1])[i] * (abs(data[i + 1, 2]) + abs(data[i, 2])) / 2
    }
    if (temp_abs_dev >= max_abs_dev) {
      max_abs_dev <- temp_abs_dev
    }
  }
  avg_abs_dev <- avg_abs_dev / nrow(data)
  c("tot_dist" = as.double(tot_dist), "max_abs_dev" = as.double(max_abs_dev),
    "avg_abs_dev" = as.double(avg_abs_dev), "AUC" = as.double(AUC))
}
result <- compute(subset)
result
as.matrix(train_measures[1,])
```

So I have got the same measures with the sample data successfully.

---

# Solution (f)

## Check my solutions against all the data from train_trajectories.

### **Relevant code**
```{r check_train}
newdatax <- matrix(0, nrow = nrow(train_measures), ncol = 6)
for (i in 1:nrow(train_measures)) {
    s_nr <- as.double(train_measures[i, 1])
    c_nr <- as.double(train_measures[i, 2])
    subset <- as.matrix(train_trajectories[train_trajectories$subject_nr == s_nr 
                                           & train_trajectories$count_trial == c_nr, 3:5])
    S <- compute(subset)
    newdatax[i, ] <- c(c("subject_nr" = s_nr, "count_trial" = c_nr), S)
}
head(newdatax,5)
head(train_measures,5)

```
So I verified my solutions against all the train samples.

---
 
## Compute for the test_trajectories

### **Relevant code**
```{r compute_test}
test_measure <- matrix(0, nrow = 5, ncol = 6)
for (i in 1:5) {
    s_nr <- 5 + i
    c_nr <- 1
    subset <- as.matrix(test_trajectories[test_trajectories$subject_nr == s_nr, 3:5])
    result_i <- compute(subset)
    test_measure[i, ] <- c(c(s_nr, c_nr), result_i)
}
# reformat the matrix into data.frame
colnames(test_measure) <- c("subject_nr", "count_trial", "tot_dist", "max_abs_dev",
                            "avg_abs_dev", "AUC")
test_measure <- data.frame(test_measure)
test_measure
```

---

# Reformat as a table

I need to report my results in a nicely formatted table:

```{r table,echo=FALSE}
knitr::kable(test_measure, caption = "The test measure results")
```







